---
title: "ONS76-SPHER_Seurat_EXPLORE1_50K"
author: "JJ"
date: "19/01/2023"
output: html_document
---

```{r graphical_output}
options(bitmapType='cairo-png') # this line of code needed BEFORE knitr code in chunk below
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(dev = "png",
                      dpi = 300,
                      echo = TRUE,   # prints the code (if FALSE then only code output is shown)
                      cache = TRUE,
                      fig.path = "./outputs50K/ONS76_SPHER/figures/")
```

```{r libraries}
suppressPackageStartupMessages({
library(Seurat)
library(glmGamPoi)
library(presto)
library(msigdbr)
library(fgsea)
library(org.Hs.eg.db)
library(DOSE)
library(clustree)
library(RColorBrewer)
library(tidyverse)
library(magrittr)
library(patchwork)
})
```

Type of analysis: clustering of the DAOY-SPHEROID sample, marker identification, cell cycle difference regression, proliferation scoring of clusters, pathway enrichment for clusters, cell cycle phase analysis. Workflow performed EXcluding mitochondrial genes, and INcluding mitochondrial genes - essentially the same results. At an early step any cells with >10% mitochondrial reads were excluded from further analyses.

***Seurat clustering of ONS76-spheroid sample***
ribosome genes retained
```{r eval=FALSE, echo=FALSE}
# load the data from scratch
# load the ONS76-2D raw data (from "working_data" folder)
data_dir <- "./working_data/ONS76-SPHER/"   # the dot is current dir and is essential!
list.files(data_dir) # should show matrix.mtx.gz, features.tsv.gz, barcodes.tsv.gz
```

```{r eval=FALSE, echo=FALSE}
# note these are .gz files which can be directly read. n = 1322 cells (o76spher.data)
o76spher.data <- ReadMtx(mtx = paste0(data_dir, "matrix.mtx.gz"),
                         features = paste0(data_dir, "features.tsv.gz"), 
                         cells = paste0(data_dir, "barcodes.tsv.gz"))
# create the seurat object, keep genes expressed in 10 or more cells
o76spher <- CreateSeuratObject(counts = o76spher.data,
                          min.cells = 10, min.genes=200,
                          project = "ons76spher")

dim(o76spher) # 17,677 genes and 1322 cells
```

```{r eval=FALSE, echo=FALSE}
# identify mitochondrial genes and calculate percent
mito.genes <- grep(pattern = "^MT", x = rownames(x = o76spher), value = TRUE)
percent.mito <- 
  Matrix::colSums(o76spher[mito.genes, ])/Matrix::colSums(o76spher)
```

```{r eval=FALSE, echo=FALSE}
# add the mitochondrial gene percent to the metadata
o76spher <- 
  AddMetaData(object = o76spher, metadata = percent.mito, col.name = "percent.mito")
```

```{r eval=FALSE, echo=FALSE}
# remove all cells with percent.mito > 10%
o76spher <- subset(o76spher, subset = percent.mito < 0.1)
dim(o76spher) # 17,664 genes and 1164 cells

# remove mitochondrial genes (this is acceptable: e.g. see https://nbisweden.github.io/excelerate-scRNAseq/session-qc/Quality_control.html)

# remove mitochondrial genes 
o76spher <- o76spher[ ! grepl("^MT-", rownames(o76spher)), ]
dim(o76spher)  # 17,664 genes, 1322 cells
```

```{r eval=FALSE, echo=FALSE}
# save the above object, non-normalised, non-clustered
saveRDS(o76spher,
        file = "./outputs50K/ONS76_SPHER/filt.MitoCellsGenes.RiboGenesRetained.seurat.ONS76spher.rds")
```

```{r eval=FALSE, echo=FALSE}
# load the above object, which has only been through mito, gene and cell filtering
o76spher <- readRDS("outputs50K/ONS76_SPHER/filt.MitoCellsGenes.RiboGenesRetained.seurat.ONS76spher.rds")
```

```{r eval=FALSE, echo=FALSE}
# load cell cycle genes, calculate cell-cycle difference in advance of regressing out CC difference

# include normalisation step of dyCbo before cell cycle scoring as per github issue: https://github.com/satijalab/seurat/issues/1679
o76spher <- NormalizeData(o76spher)

s.genes <- cc.genes$s.genes
g2m.genes <- cc.genes$g2m.genes
# calculate CC difference
o76spher <- CellCycleScoring(o76spher,
                g2m.features = g2m.genes,
                s.features = s.genes)
o76spher$CC.difference <- o76spher$S.Score - o76spher$G2M.Score
```

```{r eval=FALSE, echo=FALSE}
# this object is input for integration - already SCTransformed, CCdiff regressed, and RunPCA()
o76spher.input.integration <- SCTransform(o76spher, 
                   vars.to.regress = "CC.difference",
                   method = "glmGamPoi",
                   vst.flavor = "v2",
                   verbose = FALSE) %>% 
  RunPCA(npcs = 30, verbose = FALSE)
```

```{r eval=FALSE, echo=FALSE}
# save the SCTransformed object, CC difference regressed, PCA reduction, as input for integrated workflow
saveRDS(o76spher.input.integration, "./outputs50K/ONS76_SPHER/ONS76spher.RiboGenesRetained.inputIntegration.SCTransformed.CCdiffRegressed.PCA.unclustered.rds")
```

```{r eval=FALSE, echo=FALSE}
o76spher <- SCTransform(o76spher, vars.to.regress = "CC.difference", method = "glmGamPoi", 
                        verbose = FALSE) %>%   
  RunPCA(verbose = FALSE) %>% 
  RunUMAP(dims=1:30, verbose=FALSE) %>% 
  FindNeighbors(dims=1:30, verbose=FALSE) %>% 
  FindClusters(verbose=FALSE, resolution = 0.5)
```

```{r eval=FALSE, echo=FALSE}
# save the clustered, SCTransformed object above
saveRDS(o76spher, 
        file = "./outputs50K/ONS76_SPHER/ons76spher.RiboRetained.MitoCellsGenesFilt.SCT.clustered.CCdiffRegressed.rds")
```

```{r eval=FALSE, echo=FALSE}
# load the above object. This object not yet been annotated.
o76spher <- 
  readRDS("outputs50K/ONS76_SPHER/ons76spher.RiboRetained.MitoCellsGenesFilt.SCT.clustered.CCdiffRegressed.rds")
```

```{r eval=FALSE, echo=FALSE, DimplotONS76SspherUmapSCTransform, fig.height=6, fig.width=10}
# for res = 0.5, 5 clusters are found
DimPlot(o76spher, label = TRUE, pt.size = 1, label.size = 4)
```

SECTION END
##########################################################################################################

##########################################################################################################
***Marker identification***
```{r eval=FALSE, echo=FALSE}
# get the markers of each cluster
o76spher.POSmarkers <- FindAllMarkers(object = o76spher, only.pos = TRUE, min.pct = 0.25, logfc.threshold = 0.25)
```

```{r eval=FALSE, echo=FALSE}
saveRDS(o76spher.POSmarkers,
        file = "./outputs50K/ONS76_SPHER/POS.markers.RiboRetained.ONS76spheroid.SCT.clustered.rds")
```

```{r eval=FALSE, echo=FALSE}
o76spher.POSmarkers <- readRDS("outputs50K/ONS76_SPHER/POS.markers.RiboRetained.ONS76spheroid.SCT.clustered.rds")
```

```{r eval=FALSE, echo=FALSE}
o76spher.POSmarkers %>% 
  dplyr::group_by(cluster) %>% 
  slice_head(n = 10)
```

```{r eval=FALSE, echo=FALSE}
# MEIS1 and CNTN1 are expressed of these five, NEUROD1 is NOT expressed
o76spher.POSmarkers %>%
  filter(gene %in% c("NEUROD1", "CNTN1", "CNTN2", "MEIS1", "DCX"))
```

```{r eval=FALSE, echo=FALSE, VlnPlotIFITM3, fig.height=3, fig.width=12}
VlnPlot(o76spher, features = c("IFITM3", "SPARC"))
```


```{r eval=FALSE, echo=FALSE, VlnPlotGranuleGenesONS76spher, fig.height=17, fig.width=22}
# improed formatting of plots for manuscript
p1a <- VlnPlot(o76spher,
               features = "MEIS1") +
  theme(axis.text = element_text(size = 22),
        axis.title.x = element_blank(),
        axis.title.y = element_text(size = 22),
        legend.position = "none",
        plot.title = element_text(size = 30))

p1b <- VlnPlot(o76spher,
               features = "CNTN1") +
  theme(axis.text = element_text(size = 22),
        axis.title.x = element_blank(),
        axis.title.y = element_text(size = 22),
        legend.position = "none",
        plot.title = element_text(size = 30))

p1c <- VlnPlot(o76spher,
               features = "SATB2") +
  theme(axis.text = element_text(size = 22),
        axis.title.x = element_blank(),
        axis.title.y = element_text(size = 22),
        legend.position = "none",
        plot.title = element_text(size = 30))

p1d <- VlnPlot(o76spher,
               features = "ZIC1") +
  theme(axis.text = element_text(size = 22),
        axis.title.x = element_blank(),
        axis.title.y = element_text(size = 22),
        legend.position = "none",
        plot.title = element_text(size = 30))

p1e <- VlnPlot(o76spher,
               features = "MEIS2") +
  theme(axis.text = element_text(size = 22),
        axis.title.x = element_blank(),
        axis.title.y = element_text(size = 22),
        legend.position = "none",
        plot.title = element_text(size = 30))

p1f <- VlnPlot(o76spher,
              features = "JKAMP") +
  theme(axis.text = element_text(size = 22),
        axis.title.x = element_blank(),
        axis.title.y = element_text(size = 22),
        legend.position = "none",
        plot.title = element_text(size = 30))

p1a + p1b + p1c + p1d + p1e + p1f + plot_layout(ncol = 2)
```

```{r eval=FALSE, echo=FALSE, ViolinplotONS76spheroidCluster4ProlifernMarkers, fig.height=17, fig.width=22}
p2a <- VlnPlot(o76spher,
               features = "MKI67") +
  theme(axis.text = element_text(size = 22),
        axis.title.x = element_blank(),
        axis.title.y = element_text(size = 22),
        legend.position = "none",
        plot.title = element_text(size = 30))

p2b <- VlnPlot(o76spher,
               features = "MYBL2") +
  theme(axis.text = element_text(size = 22),
        axis.title.x = element_blank(),
        axis.title.y = element_text(size = 22),
        legend.position = "none",
        plot.title = element_text(size = 30))

p2c <- VlnPlot(o76spher,
               features = "BIRC5") +
  theme(axis.text = element_text(size = 22),
        axis.title.x = element_blank(),
        axis.title.y = element_text(size = 22),
        legend.position = "none",
        plot.title = element_text(size = 30))

p2d <- VlnPlot(o76spher,
               features = "RRM2") +
  theme(axis.text = element_text(size = 22),
        axis.title.x = element_blank(),
        axis.title.y = element_text(size = 22),
        legend.position = "none",
        plot.title = element_text(size = 30))

p2e <- VlnPlot(o76spher,
               features = "ZIC2") +
  theme(axis.text = element_text(size = 22),
        axis.title.x = element_blank(),
        axis.title.y = element_text(size = 22),
        legend.position = "none",
        plot.title = element_text(size = 30))

p2f <- VlnPlot(o76spher,
              features = "JKAMP") +
  theme(axis.text = element_text(size = 22),
        axis.title.x = element_blank(),
        axis.title.y = element_text(size = 22),
        legend.position = "none",
        plot.title = element_text(size = 30))

p2a + p2b + p2c + p2d + p2e + p2f + plot_layout(ncol = 2)

```

```{r eval=FALSE, echo=FALSE, ViolinplotONS76spheroidUniversalMarkers, fig.height=17, fig.width=22}
p3a <- VlnPlot(o76spher,
               features = "IFITM3") +
  theme(axis.text = element_text(size = 22),
        axis.title.x = element_blank(),
        axis.title.y = element_text(size = 22),
        legend.position = "none",
        plot.title = element_text(size = 30))

p3b <- VlnPlot(o76spher,
               features = "IGFBP2") +
  theme(axis.text = element_text(size = 22),
        axis.title.x = element_blank(),
        axis.title.y = element_text(size = 22),
        legend.position = "none",
        plot.title = element_text(size = 30))

p3c <- VlnPlot(o76spher,
               features = "STC1") +
  theme(axis.text = element_text(size = 22),
        axis.title.x = element_blank(),
        axis.title.y = element_text(size = 22),
        legend.position = "none",
        plot.title = element_text(size = 30))

p3d <- VlnPlot(o76spher,
               features = "SPARC") +
  theme(axis.text = element_text(size = 22),
        axis.title.x = element_blank(),
        axis.title.y = element_text(size = 22),
        legend.position = "none",
        plot.title = element_text(size = 30))

p3e <- VlnPlot(o76spher,
               features = "ZIC2") +
  theme(axis.text = element_text(size = 22),
        axis.title.x = element_blank(),
        axis.title.y = element_text(size = 22),
        legend.position = "none",
        plot.title = element_text(size = 30))

p3f <- VlnPlot(o76spher,
              features = "JKAMP") +
  theme(axis.text = element_text(size = 22),
        axis.title.x = element_blank(),
        axis.title.y = element_text(size = 22),
        legend.position = "none",
        plot.title = element_text(size = 30))

p3a + p3b + p3c + p3d + p3e + p3f + plot_layout(ncol = 2)

```

****Expression of top 10 genes of each cluster****
The 5 clusters look different from each other when ribosome genes are retained. Not much different than with ribosome genes removed. Cluster 2 and 3 are more similar to each other than to other clusters as noted previously with ribosome genes removed.
```{r eval=FALSE, echo=FALSE, HeatmapONS76spherTop10markers5clustersRes0.5, fig.height=6, fig.width=10}
# first load o76spher seurat object above and the file of positive markers, then only run this chunk
o76spher.POSmarkers %>% 
  group_by(cluster) %>%
    top_n(n = 10, wt = avg_log2FC) -> top10
DoHeatmap(o76spher, features = top10$gene) + guides(color = "none") +
  theme(text = element_text(size = 25), legend.title = element_text(size=20), 
        legend.text=element_text(size=20)) 
```


SECTION END
##########################################################################################################

##########################################################################################################
***Proliferation scoring of Seurat clusters***
```{r eval=FALSE, echo=FALSE}
s.genes <- cc.genes$s.genes
g2m.genes <- cc.genes$g2m.genes
prolif.genes <- list(c(s.genes, g2m.genes))
```


```{r eval=FALSE, echo=FALSE, AddModuleScoreProlifernONS76spheroid, fig.height=6, fig.width=10}
phase.o76spher <- AddModuleScore(o76spher,
                        features = prolif.genes,
                        name = "prolif.genes")
FeaturePlot(phase.o76spher, features = 'prolif.genes1', label = TRUE, repel = TRUE, pt.size = 1,
            label.size = 4) +
   scale_colour_gradientn(colours = rev(brewer.pal(n = 11, name = "RdBu"))) +
  labs(title = "")
```


SECTION END
##########################################################################################################

##########################################################################################################
***Functional enrichment of Seurat clusters in ONS76-spheroid sample***
Gene ontology Biological Process pathways using FGSEA

```{r eval=FALSE, echo=FALSE}
# prepare the data for gProfiler (web app) input (res = 0.5)
# use presto
o76spher.genes.presto <- wilcoxauc(o76spher, 'seurat_clusters')
nrow(o76spher.genes.presto)
dplyr::count(o76spher.genes.presto, group)
```

****Cluster 0 FGSEA****
enriched for protein translation
```{r eval=FALSE, echo=FALSE}
# filter and order the gene list for cluster 0
o76spher.genes.presto.clust0 <- o76spher.genes.presto %>% 
  dplyr::filter(group == "0" & logFC > 0.25) %>%
  dplyr::arrange(desc(auc)) %>% 
  dplyr::select(feature, auc)

nrow(o76spher.genes.presto.clust0) # 221 genes
```

```{r eval=FALSE, echo=FALSE}
# gProfiler input - use Filezilla to pull the file to my Mac and then copy and paste the ordered list
# of genes into gProfiler web app
write.csv(o76spher.genes.presto.clust0, 
          "./outputs50K/ONS76_SPHER/ONS76spheroid.RiboRetained.cluster0.genes.presto.gProfilerInput.csv", quote = F)
```

```{r eval=FALSE, echo=FALSE}
ranks.clus0 <- deframe(o76spher.genes.presto.clust0)
head(ranks.clus0)
```

```{r eval=FALSE, echo=FALSE}
# use fgsea 'multilevel' function; the multilevel function has by default NPermSimple=1000
fgseaRes.clus0.GOBP <- 
  fgseaMultilevel(fgsea_sets_GOBP, stats = ranks.clus0, eps = 0, minSize = 15, scoreType = "pos", maxSize = 200)
```

```{r eval=FALSE, echo=FALSE}
# tidy up the data
fgseaResTidy.clus0.GOBP <- fgseaRes.clus0.GOBP %>%
  as_tibble() %>%
  arrange(desc(NES))

fgseaResTidy.clus0.GOBP %>% 
  dplyr::select(-leadingEdge, -ES) %>% 
  arrange(padj) %>% 
  head(10)
```

```{r eval=FALSE, echo=FALSE}
# this is the 'tidy' version that is saved
saveRDS(fgseaResTidy.clus0.GOBP, 
        file = "./outputs50K/ONS76_SPHER/cluster0.ONS76spheroid.multifgsea.presto.GOBiolProcessGenes.rds")
```

```{r eval=FALSE, echo=FALSE}
fgseaResTidy.clus0.GOBP <- 
  readRDS("outputs50K/ONS76_SPHER/cluster0.ONS76spheroid.multifgsea.presto.GOBiolProcessGenes.rds")
```

```{r eval=FALSE, echo=FALSE}
# use fwrite() as a column of the tibble is a list!
fwrite(fgseaResTidy.clus0.GOBP, 
       file = "./outputs50K/ONS76_SPHER/cluster0.ONS76spheroid.multifgsea.presto.GOBiolProcessGenes.csv")
```


****Cluster 1 FGSEA****
not enriched for any pathways
```{r eval=FALSE, echo=FALSE}
# filter and order the gene list for cluster 1
o76spher.genes.presto.clust1 <- o76spher.genes.presto %>% 
  dplyr::filter(group == "1" & logFC > 0.25) %>%
  dplyr::arrange(desc(auc)) %>% 
  dplyr::select(feature, auc)

nrow(o76spher.genes.presto.clust1) # 385 genes
```

```{r eval=FALSE, echo=FALSE}
# gProfiler input - use Filezilla to pull the file to my Mac and then copy and paste the ordered list
# of genes into gProfiler web app
write.csv(o76spher.genes.presto.clust1, 
          "./outputs50K/ONS76_SPHER/ONS76spheroid.RiboRetained.cluster1.genes.presto.gProfilerInput.csv", quote = F)
```

```{r eval=FALSE, echo=FALSE}
ranks.clus1 <- deframe(o76spher.genes.presto.clust1)
head(ranks.clus1)
```

```{r eval=FALSE, echo=FALSE}
# use fgsea 'multilevel' function; the multilevel function has by default NPermSimple=1000
fgseaRes.clus1.GOBP <- 
  fgseaMultilevel(fgsea_sets_GOBP, stats = ranks.clus1, eps = 0, minSize = 15, scoreType = "pos", maxSize = 200)
```

```{r eval=FALSE, echo=FALSE}
# tidy up the data
fgseaResTidy.clus1.GOBP <- fgseaRes.clus1.GOBP %>%
  as_tibble() %>%
  arrange(desc(NES))

fgseaResTidy.clus1.GOBP %>% 
  dplyr::select(-leadingEdge, -ES) %>% 
  arrange(padj) %>% 
  head(10)
```

```{r eval=FALSE, echo=FALSE}
# this is the 'tidy' version that is saved - NO SIGNIFICANT ENRICHMENT OF PATHWAYS!
saveRDS(fgseaResTidy.clus1.GOBP, 
        file = "./outputs50K/ONS76_SPHER/cluster1.ONS76spheroid.multifgsea.presto.GOBiolProcessGenes.rds")
```


****Cluster 2 FGSEA****
not enriched for any pathways
```{r eval=FALSE, echo=FALSE}
# filter and order the gene list for cluster 2
o76spher.genes.presto.clust2 <- o76spher.genes.presto %>% 
  dplyr::filter(group == "2" & logFC > 0.25) %>%
  dplyr::arrange(desc(auc)) %>% 
  dplyr::select(feature, auc)

nrow(o76spher.genes.presto.clust2) # only 69 genes
```

```{r eval=FALSE, echo=FALSE}
# gProfiler input - use Filezilla to pull the file to my Mac and then copy and paste the ordered list
# of genes into gProfiler web app
write.csv(o76spher.genes.presto.clust2, 
          "./outputs50K/ONS76_SPHER/ONS76spheroid.RiboRetained.cluster2.genes.presto.gProfilerInput.csv", quote = F)
```

```{r eval=FALSE, echo=FALSE}
ranks.clus2 <- deframe(o76spher.genes.presto.clust2)
head(ranks.clus2)
```

```{r eval=FALSE, echo=FALSE}
# use fgsea 'multilevel' function; the multilevel function has by default NPermSimple=1000
fgseaRes.clus2.GOBP <- 
  fgseaMultilevel(fgsea_sets_GOBP, stats = ranks.clus2, eps = 0, minSize = 15, scoreType = "pos", maxSize = 200)
```

```{r eval=FALSE, echo=FALSE}
# tidy up the data
fgseaResTidy.clus2.GOBP <- fgseaRes.clus2.GOBP %>%
  as_tibble() %>%
  arrange(desc(NES))

fgseaResTidy.clus2.GOBP %>% 
  dplyr::select(-leadingEdge, -ES) %>% 
  arrange(padj) %>% 
head(20)   # no significant pathway enrichment!
```

```{r eval=FALSE, echo=FALSE}
# this is the 'tidy' version that is saved - NO SIGNIFICANT PATHWAY ENRICHMENT!
saveRDS(fgseaResTidy.clus2.GOBP, 
        file = "./outputs50K/ONS76_SPHER/cluster2.ONS76spheroid.multifgsea.presto.GOBiolProcessGenes.rds")
```


****Cluster 3 FGSEA****
no functional enrichment
```{r eval=FALSE, echo=FALSE}
# filter and order the gene list for cluster 3
o76spher.genes.presto.clust3 <- o76spher.genes.presto %>% 
  dplyr::filter(group == "3" & logFC > 0.25) %>%
  dplyr::arrange(desc(auc)) %>% 
  dplyr::select(feature, auc)

nrow(o76spher.genes.presto.clust3) # 238 genes
```

```{r eval=FALSE, echo=FALSE}
# gProfiler input - use Filezilla to pull the file to my Mac and then copy and paste the ordered list
# of genes into gProfiler web app
write.csv(o76spher.genes.presto.clust3, 
          "./outputs50K/ONS76_SPHER/ONS76spheroid.RiboRetained.cluster3.genes.presto.gProfilerInput.csv", quote = F)
```

```{r eval=FALSE, echo=FALSE}
ranks.clus3 <- deframe(o76spher.genes.presto.clust3)
head(ranks.clus3)
```

```{r eval=FALSE, echo=FALSE}
# use fgsea 'multilevel' function; the multilevel function has by default NPermSimple=1000
fgseaRes.clus3.GOBP <- 
  fgseaMultilevel(fgsea_sets_GOBP, stats = ranks.clus3, eps = 0, minSize = 15, scoreType = "pos", maxSize = 200)
```

```{r eval=FALSE, echo=FALSE}
# tidy up the data
fgseaResTidy.clus3.GOBP <- fgseaRes.clus3.GOBP %>%
  as_tibble() %>%
  arrange(desc(NES))

fgseaResTidy.clus3.GOBP %>% 
  dplyr::select(-leadingEdge, -ES) %>% 
  arrange(padj) %>% 
  head(10) # no significant pathway enrichment! (using padj < 0.05)
```

```{r eval=FALSE, echo=FALSE}
# this is the 'tidy' version that is saved - NO SIGNIFICANT PATHWAY ENRICHMENT!
saveRDS(fgseaResTidy.clus3.GOBP, 
        file = "./outputs50K/ONS76_SPHER/cluster3.ONS76spheroid.multifgsea.presto.GOBiolProcessGenes.rds")
```


****Cluster 4 FGSEA****
enriched for cell cycle/proliferation genes
```{r eval=FALSE, echo=FALSE}
# filter and order the gene list for cluster 4
o76spher.genes.presto.clust4 <- o76spher.genes.presto %>% 
  dplyr::filter(group == "4" & logFC > 0.25) %>%
  dplyr::arrange(desc(auc)) %>% 
  dplyr::select(feature, auc)

nrow(o76spher.genes.presto.clust4) # 588 genes
```

```{r eval=FALSE, echo=FALSE}
ranks.clus4 <- deframe(o76spher.genes.presto.clust4)
head(ranks.clus4)
```

```{r eval=FALSE, echo=FALSE}
# use fgsea 'multilevel' function; the multilevel function has by default NPermSimple=1000
fgseaRes.clus4.GOBP <- 
  fgseaMultilevel(fgsea_sets_GOBP, stats = ranks.clus4, eps = 0, minSize = 15, scoreType = "pos", maxSize = 200)
```

```{r eval=FALSE, echo=FALSE}
# tidy up the data - enrichment for cell division/proliferation
fgseaResTidy.clus4.GOBP <- fgseaRes.clus4.GOBP %>%
  as_tibble() %>%
  arrange(desc(NES))

fgseaResTidy.clus4.GOBP %>% 
  dplyr::select(-leadingEdge, -ES) %>% 
  arrange(padj) %>% 
  head(10) 
```

```{r eval=FALSE, echo=FALSE}
# this is the 'tidy' version that is saved
saveRDS(fgseaResTidy.clus4.GOBP, 
        file = "./outputs50K/ONS76_SPHER/cluster4.ONS76spheroid.multifgsea.presto.GOBiolProcessGenes.rds")
```

```{r eval=FALSE, echo=FALSE}
# use fwrite() as a column of the tibble is a list!
fwrite(fgseaResTidy.clus4.GOBP, 
       file = "./outputs50K/ONS76_SPHER/cluster4.ONS76spheroid.multifgsea.presto.GOBiolProcessGenes.csv")
```


```{r eval=FALSE, echo=FALSE}
# gProfiler input - use Filezilla to pull the file to my Mac and then copy and paste the ordered list
# of genes into gProfiler web app
write.csv(o76spher.genes.presto.clust4, 
          "./outputs50K/ONS76_SPHER/ONS76spheroid.RiboRetained.cluster4.genes.presto.gProfilerInput.csv", quote = F)
```

SECTION END
##########################################################################################################

##########################################################################################################
***Cell cycle phase analysis***
```{r eval=FALSE, echo=FALSE}
# get the number of cells per cluster
table(Idents(o76spher))
```

```{r eval=FALSE, echo=FALSE}
# load the dataset
o76spher <- readRDS("outputs50K/ONS76_SPHER/ons76spher.RiboRetained.MitoCellsGenesFilt.SCT.clustered.CCdiffRegressed.rds")
```

```{r eval=FALSE, echo=FALSE}
# confirm that cell cycle phase metadata is present - it is
head(o76spher@meta.data)
```

```{r eval=FALSE, echo=FALSE}
# extract the G1 only cells from o76spher (this 'o76spher' object is the already CLUSTERED object)
# first, switch Idents to 'Phase'
Idents(o76spher) <- "Phase"
o76spher.g1.cellIDs <- WhichCells(o76spher, idents = "G1")
length(o76spher.g1.cellIDs) # 635 cells
```

```{r eval=FALSE, echo=FALSE}
# these cell IDs look correctly formatted
# object o76spher has to be "filt.RiboMitoCellsGenes.seurat.ons76spheroid.rds" above!
o76spher.g1cells <- subset(o76spher, cells = o76spher.g1.cellIDs)
head(WhichCells(o76spher.g1cells))
```

```{r eval=FALSE, echo=FALSE}
# check the number of cells is correct in the subsetted object
dim(o76spher.g1cells) # 635 cells in G1
```


```{r eval=FALSE, echo=FALSE}
# create an object suitable for integration (don't need to regress cell cycle!)

o76spher.g1cells.input.integration <- SCTransform(o76spher.g1cells,
                                                  method = "glmGamPoi",
                                                  vst.flavor = "v2",
                                                  verbose = FALSE) %>% 
  RunPCA(npcs = 30, verbose = FALSE)
```

```{r eval=FALSE, echo=FALSE}
dim(o76spher.g1cells.input.integration) # object has 635 cells which is correct
```

```{r eval=FALSE, echo=FALSE}
# save the "o76spher.g1cells.input.integration" object
saveRDS(o76spher.g1cells.input.integration, 
        file = "./outputs50K/ONS76_SPHER/G1phaseCells.RiboRetained.ONS76spheroid.inputIntegration.SCTransformed.PCA.unclustered.rds")
```

SECTION END
##########################################################################################################

##########################################################################################################
##########################################################################################################
***Alternate workflow with removal of cells with > 10% mitochondrial gene expression, RETENTION of mitochondrial genes (WITHOUT regression)***
basically, the same result as following removal of mitochondrial genes.
```{r eval=FALSE, echo=FALSE}
# load the data from scratch
# load the ONS76-2D raw data (from "working_data" folder)
data_dir <- "./working_data/ONS76-SPHER/"   # the dot is current dir and is essential!
list.files(data_dir) # should show matrix.mtx.gz, features.tsv.gz, barcodes.tsv.gz
```

```{r eval=FALSE, echo=FALSE}
# note these are .gz files which can be directly read. n = 1322 cells (o76spher.data)
o76spher.data <- ReadMtx(mtx = paste0(data_dir, "matrix.mtx.gz"),
                         features = paste0(data_dir, "features.tsv.gz"), 
                         cells = paste0(data_dir, "barcodes.tsv.gz"))
# create the seurat object, keep genes expressed in 10 or more cells
o76spher <- CreateSeuratObject(counts = o76spher.data,
                          min.cells = 10, min.genes=200,
                          project = "ons76spher")

dim(o76spher) # 17,677 genes and 1322 cells
```

```{r eval=FALSE, echo=FALSE}
# identify mitochondrial genes and calculate percent
mito.genes <- grep(pattern = "^MT", x = rownames(x = o76spher), value = TRUE)
percent.mito <- 
  Matrix::colSums(o76spher[mito.genes, ])/Matrix::colSums(o76spher)
```

```{r eval=FALSE, echo=FALSE}
# add the mitochondrial gene percent to the metadata
o76spher <- 
  AddMetaData(object = o76spher, metadata = percent.mito, col.name = "percent.mito")
```

```{r eval=FALSE, echo=FALSE}
# remove all cells with percent.mito > 10%, but do NOT remove mitochondrial genes, nor regress out mitochondrial genes
o76spher <- subset(o76spher, subset = percent.mito < 0.1)
dim(o76spher) # 17,664 genes and 1164 cells
```

```{r eval=FALSE, echo=FALSE}
# save the above object, non-normalised, non-clustered
saveRDS(o76spher,
        file = "./outputs50K/ONS76_SPHER/MITO_RETAIN/filt.MitoCellsGenes.RiboGenesRetained.MitoGenesRetainedUNregressed.seurat.ONS76spher.rds")
```

```{r eval=FALSE, echo=FALSE}
# load cell cycle genes, calculate cell-cycle difference in advance of regressing out CC difference

# include normalisation step of dyCbo before cell cycle scoring as per github issue: https://github.com/satijalab/seurat/issues/1679
o76spher <- NormalizeData(o76spher)

s.genes <- cc.genes$s.genes
g2m.genes <- cc.genes$g2m.genes
# calculate CC difference
o76spher <- CellCycleScoring(o76spher,
                g2m.features = g2m.genes,
                s.features = s.genes)
o76spher$CC.difference <- o76spher$S.Score - o76spher$G2M.Score
```

```{r eval=FALSE, echo=FALSE}
# this object is input for integration - already SCTransformed, CCdiff regressed, and RunPCA()
o76spher.input.integration <- SCTransform(o76spher, 
                   vars.to.regress = "CC.difference",
                   method = "glmGamPoi",
                   vst.flavor = "v2",
                   verbose = FALSE) %>% 
  RunPCA(npcs = 30, verbose = FALSE)
```

```{r eval=FALSE, echo=FALSE}
# save the SCTransformed object, CC difference regressed, PCA reduction, as input for integrated workflow
saveRDS(o76spher.input.integration, "./outputs50K/ONS76_SPHER/MITO_RETAIN/ONS76spher.RiboGenesRetained.MitoRetainedUNregressed.inputIntegration.SCTransformed.CCdiffRegressed.PCA.unclustered.rds")
```

```{r eval=FALSE, echo=FALSE}
o76spher <- SCTransform(o76spher, vars.to.regress = "CC.difference", method = "glmGamPoi", 
                        vst.flavor = "v2", verbose = FALSE) %>%   
  RunPCA(verbose = FALSE) %>% 
  RunUMAP(dims=1:30, verbose=FALSE) %>% 
  FindNeighbors(dims=1:30, verbose=FALSE) %>% 
  FindClusters(verbose=FALSE, resolution = 0.5)
```

```{r eval=FALSE, echo=FALSE}
# save the clustered, SCTransformed object above
saveRDS(o76spher, 
        file = "./outputs50K/ONS76_SPHER/MITO_RETAIN/ons76spher.RiboRetained.MitoRetainedUNregressed.CellsGenesFilt.SCT.clustered.CCdiffRegressed.rds")
```

```{r eval=FALSE, echo=FALSE, DimplotONS76SspherUmapSCTransform, fig.height=6, fig.width=10}
# for res = 0.5, 5 clusters are found
DimPlot(o76spher, label = TRUE, pt.size = 1, label.size = 4)
```

```{r eval=FALSE, echo=FALSE}
# get the markers of each cluster
o76spher.POSmarkers <- FindAllMarkers(object = o76spher, only.pos = TRUE, min.pct = 0.25, logfc.threshold = 0.25)
```

```{r eval=FALSE, echo=FALSE}
o76spher.POSmarkers %>% 
  dplyr::group_by(cluster) %>% 
  slice_head(n = 20)
```

```{r eval=FALSE, echo=FALSE}
# MEIS1 and CNTN1 are expressed of these five, but NEUROD1 is NOT expressed
o76spher.POSmarkers %>%
  filter(gene %in% c("NEUROD1", "CNTN1", "CNTN2", "MEIS1", "DCX"))
```

****Proliferation scoring of Seurat clusters****
```{r eval=FALSE, echo=FALSE}
s.genes <- cc.genes$s.genes
g2m.genes <- cc.genes$g2m.genes
prolif.genes <- list(c(s.genes, g2m.genes))
```


```{r eval=FALSE, echo=FALSE, AddModuleScoreProlifernONS76spheroid, fig.height=6, fig.width=10}
phase.o76spher <- AddModuleScore(o76spher,
                        features = prolif.genes,
                        name = "prolif.genes")
FeaturePlot(phase.o76spher, features = 'prolif.genes1', label = TRUE, repel = TRUE, pt.size = 1,
            label.size = 4) +
   scale_colour_gradientn(colours = rev(brewer.pal(n = 11, name = "RdBu"))) +
  labs(title = "")
```

****Functional enrichment of Seurat clusters in ONS76-spheroid sample****
Gene ontology Biological Process pathways using FGSEA
-basically the same enrichment as with exclusion of mitochondrial genes: protein synthesis and proliferation
```{r eval=FALSE, echo=FALSE}
# prepare the data for gProfiler (web app) input (res = 0.5)
# use presto
o76spher.genes.presto <- wilcoxauc(o76spher, 'seurat_clusters')
nrow(o76spher.genes.presto)
dplyr::count(o76spher.genes.presto, group)
```

```{r eval=FALSE, echo=FALSE}
# use dplyr to get the appropriate gene set
msig.hs <- msigdbr(species = "Homo sapiens")

# GO biological process (GO:BP)
msig.GOBP.hs <- msig.hs %>% 
dplyr::filter(gs_subcat == "GO:BP")

head(msig.GOBP.hs)
```

```{r eval=FALSE, echo=FALSE}
fgsea_sets_GOBP <- msig.GOBP.hs %>% split(x = .$gene_symbol, f = .$gs_name)
```
*****Cluster 0 FGSEA*****
enriched for protein translation/synthesis
```{r eval=FALSE, echo=FALSE}
# filter and order the gene list for cluster 0
o76spher.genes.presto.clust0 <- o76spher.genes.presto %>% 
  dplyr::filter(group == "0" & logFC > 0.25) %>%
  dplyr::arrange(desc(auc)) %>% 
  dplyr::select(feature, auc)

nrow(o76spher.genes.presto.clust0) # 229 genes
```

```{r eval=FALSE, echo=FALSE}
# gProfiler input - use Filezilla to pull the file to my Mac and then copy and paste the ordered list
# of genes into gProfiler web app
write.csv(o76spher.genes.presto.clust0, 
          "./outputs50K/ONS76_SPHER/MITO_RETAIN/ONS76spheroid.RiboRetained.MitoRetainedUNregressed.cluster0.genes.presto.gProfilerInput.csv", quote = F)
```

```{r eval=FALSE, echo=FALSE}
ranks.clus0 <- deframe(o76spher.genes.presto.clust0)
head(ranks.clus0)
```

```{r eval=FALSE, echo=FALSE}
# use fgsea 'multilevel' function; the multilevel function has by default NPermSimple=1000
fgseaRes.clus0.GOBP <- 
  fgseaMultilevel(fgsea_sets_GOBP, stats = ranks.clus0, eps = 0, minSize = 15, scoreType = "pos", maxSize = 200)
```

```{r eval=FALSE, echo=FALSE}
# tidy up the data
fgseaResTidy.clus0.GOBP <- fgseaRes.clus0.GOBP %>%
  as_tibble() %>%
  arrange(desc(NES))

fgseaResTidy.clus0.GOBP %>% 
  dplyr::select(-leadingEdge, -ES) %>% 
  arrange(padj) %>% 
  head(10)
```

*****Cluster 1 FGSEA*****
not enriched for any pathways
```{r eval=FALSE, echo=FALSE}
# filter and order the gene list for cluster 1
o76spher.genes.presto.clust1 <- o76spher.genes.presto %>% 
  dplyr::filter(group == "1" & logFC > 0.25) %>%
  dplyr::arrange(desc(auc)) %>% 
  dplyr::select(feature, auc)

nrow(o76spher.genes.presto.clust1) # 362 genes
```

```{r eval=FALSE, echo=FALSE}
# gProfiler input - use Filezilla to pull the file to my Mac and then copy and paste the ordered list
# of genes into gProfiler web app
write.csv(o76spher.genes.presto.clust1, 
          "./outputs50K/ONS76_SPHER/MITO_RETAIN/ONS76spheroid.RiboRetained.MitoRetainUNregressed.cluster1.genes.presto.gProfilerInput.csv", quote = F)
```

```{r eval=FALSE, echo=FALSE}
ranks.clus1 <- deframe(o76spher.genes.presto.clust1)
head(ranks.clus1)
```

```{r eval=FALSE, echo=FALSE}
# use fgsea 'multilevel' function; the multilevel function has by default NPermSimple=1000
fgseaRes.clus1.GOBP <- 
  fgseaMultilevel(fgsea_sets_GOBP, stats = ranks.clus1, eps = 0, minSize = 15, scoreType = "pos", maxSize = 200)
```

```{r eval=FALSE, echo=FALSE}
# tidy up the data
fgseaResTidy.clus1.GOBP <- fgseaRes.clus1.GOBP %>%
  as_tibble() %>%
  arrange(desc(NES))

fgseaResTidy.clus1.GOBP %>% 
  dplyr::select(-leadingEdge, -ES) %>% 
  arrange(padj) %>% 
  head(10)
```

```{r eval=FALSE, echo=FALSE}
# this is the 'tidy' version that is saved - NO SIGNIFICANT ENRICHMENT OF PATHWAYS!
saveRDS(fgseaResTidy.clus1.GOBP, 
        file = "./outputs50K/ONS76_SPHER/MITO_RETAIN/cluster1.ONS76spheroid.MitoRetainUNregressed.multifgsea.presto.GOBiolProcessGenes.rds")
```


*****Cluster 2 FGSEA*****
not enriched for any pathways
```{r eval=FALSE, echo=FALSE}
# filter and order the gene list for cluster 2
o76spher.genes.presto.clust2 <- o76spher.genes.presto %>% 
  dplyr::filter(group == "2" & logFC > 0.25) %>%
  dplyr::arrange(desc(auc)) %>% 
  dplyr::select(feature, auc)

nrow(o76spher.genes.presto.clust2) # only 70 genes
```

```{r eval=FALSE, echo=FALSE}
# gProfiler input - use Filezilla to pull the file to my Mac and then copy and paste the ordered list
# of genes into gProfiler web app
write.csv(o76spher.genes.presto.clust2, 
          "./outputs50K/ONS76_SPHER/MITO_RETAIN/ONS76spheroid.RiboRetained.MitoRetainUNregressed.cluster2.genes.presto.gProfilerInput.csv", quote = F)
```

```{r eval=FALSE, echo=FALSE}
ranks.clus2 <- deframe(o76spher.genes.presto.clust2)
head(ranks.clus2)
```

```{r eval=FALSE, echo=FALSE}
# use fgsea 'multilevel' function; the multilevel function has by default NPermSimple=1000
fgseaRes.clus2.GOBP <- 
  fgseaMultilevel(fgsea_sets_GOBP, stats = ranks.clus2, eps = 0, minSize = 15, scoreType = "pos", maxSize = 200)
```

```{r eval=FALSE, echo=FALSE}
# tidy up the data
fgseaResTidy.clus2.GOBP <- fgseaRes.clus2.GOBP %>%
  as_tibble() %>%
  arrange(desc(NES))

fgseaResTidy.clus2.GOBP %>% 
  dplyr::select(-leadingEdge, -ES) %>% 
  arrange(padj) %>% 
head(20)   # no significant pathway enrichment!
```

```{r eval=FALSE, echo=FALSE}
# this is the 'tidy' version that is saved - NO SIGNIFICANT PATHWAY ENRICHMENT!
saveRDS(fgseaResTidy.clus2.GOBP, 
        file = "./outputs50K/ONS76_SPHER/cluster2.ONS76spheroid.multifgsea.presto.GOBiolProcessGenes.rds")
```


*****Cluster 3 FGSEA*****
no functional enrichment
```{r eval=FALSE, echo=FALSE}
# filter and order the gene list for cluster 3
o76spher.genes.presto.clust3 <- o76spher.genes.presto %>% 
  dplyr::filter(group == "3" & logFC > 0.25) %>%
  dplyr::arrange(desc(auc)) %>% 
  dplyr::select(feature, auc)

nrow(o76spher.genes.presto.clust3) # 212 genes
```

```{r eval=FALSE, echo=FALSE}
# gProfiler input - use Filezilla to pull the file to my Mac and then copy and paste the ordered list
# of genes into gProfiler web app
write.csv(o76spher.genes.presto.clust3, 
          "./outputs50K/ONS76_SPHER/MITO_RETAIN/ONS76spheroid.RiboRetained.MitoRetainUNregressed.cluster3.genes.presto.gProfilerInput.csv", quote = F)
```

```{r eval=FALSE, echo=FALSE}
ranks.clus3 <- deframe(o76spher.genes.presto.clust3)
head(ranks.clus3)
```

```{r eval=FALSE, echo=FALSE}
# use fgsea 'multilevel' function; the multilevel function has by default NPermSimple=1000
fgseaRes.clus3.GOBP <- 
  fgseaMultilevel(fgsea_sets_GOBP, stats = ranks.clus3, eps = 0, minSize = 15, scoreType = "pos", maxSize = 200)
```

```{r eval=FALSE, echo=FALSE}
# tidy up the data
fgseaResTidy.clus3.GOBP <- fgseaRes.clus3.GOBP %>%
  as_tibble() %>%
  arrange(desc(NES))

fgseaResTidy.clus3.GOBP %>% 
  dplyr::select(-leadingEdge, -ES) %>% 
  arrange(padj) %>% 
  head(10) # no significant pathway enrichment! (using padj < 0.05)
```

```{r eval=FALSE, echo=FALSE}
# this is the 'tidy' version that is saved - NO SIGNIFICANT PATHWAY ENRICHMENT!
saveRDS(fgseaResTidy.clus3.GOBP, 
        file = "./outputs50K/ONS76_SPHER/MITO_RETAIN/cluster3.ONS76spheroid.MitoRetainUNregressed.multifgsea.presto.GOBiolProcessGenes.rds")
```


*****Cluster 4 FGSEA*****
enriched for macromolecule biosynthesis/peptide biosnthesis
```{r eval=FALSE, echo=FALSE}
# filter and order the gene list for cluster 4
o76spher.genes.presto.clust4 <- o76spher.genes.presto %>% 
  dplyr::filter(group == "4" & logFC > 0.25) %>%
  dplyr::arrange(desc(auc)) %>% 
  dplyr::select(feature, auc)

nrow(o76spher.genes.presto.clust4) # 378 genes
```

```{r eval=FALSE, echo=FALSE}
ranks.clus4 <- deframe(o76spher.genes.presto.clust4)
head(ranks.clus4)
```

```{r eval=FALSE, echo=FALSE}
# use fgsea 'multilevel' function; the multilevel function has by default NPermSimple=1000
fgseaRes.clus4.GOBP <- 
  fgseaMultilevel(fgsea_sets_GOBP, stats = ranks.clus4, eps = 0, minSize = 15, scoreType = "pos", maxSize = 200)
```

```{r eval=FALSE, echo=FALSE}
# tidy up the data - enrichment for cell division/proliferation
fgseaResTidy.clus4.GOBP <- fgseaRes.clus4.GOBP %>%
  as_tibble() %>%
  arrange(desc(NES))

fgseaResTidy.clus4.GOBP %>% 
  dplyr::select(-leadingEdge, -ES) %>% 
  arrange(padj) %>% 
  head(10) 
```

```{r eval=FALSE, echo=FALSE}
# this is the 'tidy' version that is saved
saveRDS(fgseaResTidy.clus4.GOBP, 
        file = "./outputs50K/ONS76_SPHER/MITO_RETAIN/cluster4.ONS76spheroid.multifgsea.MitoRetainUNregressed.presto.GOBiolProcessGenes.rds")
```

```{r eval=FALSE, echo=FALSE}
# use fwrite() as a column of the tibble is a list!
fwrite(fgseaResTidy.clus4.GOBP, 
       file = "./outputs50K/ONS76_SPHER/MITO_RETAIN/cluster4.ONS76spheroid.multifgsea.MitoRetainUNregressed.presto.GOBiolProcessGenes.csv")
```


```{r eval=FALSE, echo=FALSE}
# gProfiler input - use Filezilla to pull the file to my Mac and then copy and paste the ordered list
# of genes into gProfiler web app
write.csv(o76spher.genes.presto.clust4, 
          "./outputs50K/ONS76_SPHER/MITO_RETAIN/ONS76spheroid.RiboRetained.MitoRetainUNregressed.cluster4.genes.presto.gProfilerInput.csv", quote = F)
```

*****Cluster 5 FGSEA*****
enriched for cell proliferation/cell cycle pathways
```{r eval=FALSE, echo=FALSE}
# filter and order the gene list for cluster 4
o76spher.genes.presto.clust5 <- o76spher.genes.presto %>% 
  dplyr::filter(group == "5" & logFC > 0.25) %>%
  dplyr::arrange(desc(auc)) %>% 
  dplyr::select(feature, auc)

nrow(o76spher.genes.presto.clust5) # 560 genes
```

```{r eval=FALSE, echo=FALSE}
ranks.clus5 <- deframe(o76spher.genes.presto.clust5)
head(ranks.clus5)
```

```{r eval=FALSE, echo=FALSE}
# use fgsea 'multilevel' function; the multilevel function has by default NPermSimple=1000
fgseaRes.clus5.GOBP <- 
  fgseaMultilevel(fgsea_sets_GOBP, stats = ranks.clus5, eps = 0, minSize = 15, scoreType = "pos", maxSize = 200)
```

```{r eval=FALSE, echo=FALSE}
# tidy up the data - enrichment for cell division/proliferation
fgseaResTidy.clus5.GOBP <- fgseaRes.clus5.GOBP %>%
  as_tibble() %>%
  arrange(desc(NES))

fgseaResTidy.clus5.GOBP %>% 
  dplyr::select(-leadingEdge, -ES) %>% 
  arrange(padj) %>% 
  head(10) 
```

```{r eval=FALSE, echo=FALSE}
# this is the 'tidy' version that is saved
saveRDS(fgseaResTidy.clus5.GOBP, 
        file = "./outputs50K/ONS76_SPHER/MITO_RETAIN/cluster5.ONS76spheroid.multifgsea.MitoRetainUNregressed.presto.GOBiolProcessGenes.rds")
```

```{r eval=FALSE, echo=FALSE}
# use fwrite() as a column of the tibble is a list!
fwrite(fgseaResTidy.clus5.GOBP, 
       file = "./outputs50K/ONS76_SPHER/MITO_RETAIN/cluster5.ONS76spheroid.multifgsea.MitoRetainUNregressed.presto.GOBiolProcessGenes.csv")
```


```{r eval=FALSE, echo=FALSE}
# gProfiler input - use Filezilla to pull the file to my Mac and then copy and paste the ordered list
# of genes into gProfiler web app
write.csv(o76spher.genes.presto.clust5, 
          "./outputs50K/ONS76_SPHER/MITO_RETAIN/ONS76spheroid.RiboRetained.MitoRetainUNregressed.cluster5.genes.presto.gProfilerInput.csv", quote = F)
```

```{r eval=FALSE, echo=FALSE}
sessionInfo()
```


SECTION END
##########################################################################################################
